---
title: "Model for classification of human activities"
output: html_document
---

## Introduction 

This project provides a method to predict people's activities during exercise.
The training and testing data come from this source: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har).

## Data Processing

1. Downloading data sets
```{r echo =TRUE}
library(RCurl)
library(caret)
library(randomForest)
if(!file.exists("data")){dir.create("data")}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="./data/pm1train.csv", method = "curl")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="./pm1test.csv", method="curl")
train <- read.csv("./data/pm1train.csv",na.strings=c("NA", ""," "))
test <- read.csv("./data/pm1test.csv",na.strings=c("NA","", " "))
dim(train); dim(test)
```

Notice that there are 160 columns, further examination with summary() function shows that many of these columns contain significant amount of missing values, therefore, a cleaning step is necessary.

2. Cleaning training and testing data

Besides removing missing values, there are columns for identity information, such as index, names etc, those values are irrelevant and can be removed.
```{r echo=TRUE}
rmna <- apply(train, 2, function(x) sum(is.na(x)))
traincl <- train[,(rmna==0)]
testcl <- test[,(rmna==0)]
dim(traincl); dim(testcl)
# furthermore, remove lable columns, such as names, index, which are columns 1~7
traincl <- traincl[,8:60]
testcl <- testcl[,8:60]
```

## Model building and evaluation

A random forest method is used to predict the activities classes. 
```{r echo=TRUE}
model <- randomForest(classe~., data=traincl)
model
```

From the fitting data, the OOB error rate is 0.28%, in other words, the accuracy is 1-0.28%= 99.72%. You can also get the accuracy manually by adding up the numbers on the diagonal, which are correctly calculated by the method. 

Since I use a random forest method, cross validation is no needed for this algorithm. Here is a very detailed link about [random forests](https://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr) in case you are interested.  

## Predicting with test data

Use the model built to predict on the test data. 
```{r echo=TRUE}
predtest <- predict(model, testcl)
predtest
```

## Conclusion
A random forest model is used to predict the activities classes for the 20 test data. The training data shows an accuracy of 99.72%.




